### 2017-12-02 reports/poster-agu-2017/blocks/at-a-glance.tex
update poster
possible move to final poster version

\begin{block}{At a Glance}
  Beginning in Nov. 2015, repeated intense rainfall events associated with mesoscale convective activity caused severe flooding along the Paraguay-Paran\'{a} river system (\cref{fig:study-area}, red box), \emph{displacing over 170000 people}.
  We use a weather typing approch within a diagnostic approach to show that:
  \begin{itemize}
    \item Moisture and energy advection via the South American Low-Level Jet \cite{Marengo:2012cm}, particularly during ``No-Chaco'' events \cite{Vera:2006ib} favored mesoscale convective activity
    \item Strong El Ni\~{n}o favored strong SALLJ but MJO and Atlantic dipole pattern contributed to jet exit region (thus No-Chaco SALLJ events)
    \item Raw S2S model forecasts of rainfall had limited skill beyond 10-15 days, but statistical methods -- particularly the PCR and CCA methods that correct both spatial patterns and magnitudes -- substantially improve forecast skill
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.45\textwidth]{asuncion-inundaciones.jpg}~
    \noindent\includegraphics[width=0.45\textwidth]{rio-py-banados.jpg}
  	\caption{
  		Asuncion, Paraguay, 2015-16
  	}
    \label{fig:floods}
  \end{figure}
  \end{mdframed}
\end{block}

>>>>

\begin{block}{At a Glance}
  Beginning in Nov. 2015, repeated intense rainfall events associated with mesoscale convective activity caused severe flooding along the Paraguay-Paran\'{a} river system (\cref{fig:study-area}, red box), \emph{displacing over 170000 people}.
  We use a weather typing approach within a diagnostic framework to show that:
  \begin{itemize}
    \item Moisture and energy advection via the South American Low-Level Jet \cite[SALLJ;][]{Marengo:2012cm}, particularly during ``No-Chaco'' events \cite{Vera:2006ib} favored mesoscale convective activity
    \item Numerical forecasts predicted enhanced risk of heavy rainfall at the seasonal scale, but biases in spatial patterns suggest difficulties representing Pacific-Atlantic interaction
    \item Uncorrected sub-seasonal model forecasts of rainfall had limited skill beyond 10-15 days; use of Model Output Statistics -- particularly methods that correct both spatial patterns and magnitudes -- substantially improved forecast skill
    \item Strong El Ni\~{n}o and active MJO favored a strong SALLJ but an Atlantic dipole pattern influenced the jet's exit region
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.45\textwidth]{asuncion-inundaciones.jpg}~
    \noindent\includegraphics[width=0.45\textwidth]{rio-py-banados.jpg}
  	\caption{
  		Asuncion, Paraguay, 2015-16
  	}
    \label{fig:floods}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-02 reports/poster-agu-2017/blocks/obs-1516.tex
update poster
possible move to final poster version

  During NDJF 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
  Monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
  In January 2016 it weakened before returning in February 2016.
  The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.
  Weather types 1 and 4  occurred more frequency than their climatology.

  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{circulation-NDJF-1516-anomaly-alt.pdf}
  	\caption{
      	Monthly composite anomalies observed during NDJF 2015-16.
        Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
        Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
        \label{fig:anomalies}
  	}
  \end{figure}
  \end{mdframed}
\end{block}

>>>>

  During austral summer (NDJF) 2015-16, the most heavy rainfall occurred during weather types 1 and 4 (\cref{fig:rain-wt}).
  Monthly-scale circulation anomalies (\cref{fig:anomalies}) show a weak anticyclonic circulation that set up over central Brazil during November 2015 and strengthened into the following month.
  In January 2016 it weakened before returning in February 2016.
  The observed rainfall and circulation anomalies are consistent with the aggregation of the observed weather types shown in \cref{fig:rain-wt} and discussed above.

  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{circulation-NDJF-1516-anomaly-alt.pdf}
  	\caption{
      	Monthly composite anomalies observed during NDJF 2015-16.
        Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
        Contours: anomalies of \SI{850}{\hecto\pascal} streamfunction [contour interval \SI{1e4}{\meter\squared\per\second}]
        \label{fig:anomalies}
  	}
  \end{figure}
  \end{mdframed}
\end{block}

### 2017-12-02 reports/poster-agu-2017/blocks/obs-weather-types.tex
update poster
possible move to final poster version

\begin{block}{Observations and Weather Types}
  Observations come from:
  \begin{itemize}
    \item Rainfall: CPC Global Unified \cite{xie2010cpc}
    \item Atmosphere: NCAR-NCEP Reanalysis II \cite{Kanamitsu:2002kk}
  \end{itemize}
  We use weather typing \cite{Munoz2015} to identify daily circulation patterns:
  \begin{enumerate}
    \item Calculate streamfunction $\Psi$ from meridional and zonal wind \cite{Dawson:2016ge}
    \item Project \SI{850}{\hecto\pascal} streamfunction onto leading 4 EOFs
    \item $K$-means clustering using classifiability index \cite{Michelangeli1995} to generate single weather type for each day
  \end{enumerate}
  Weather typing simplifies dynamics of daily rainfall but \emph{facilitates analysis of sequences of daily weather patterns}.
  They are associated with patterns that have been well described in the literature; particularly relevant are:
  \begin{itemize}
    \item WT1 represents ``Chaco'' jet event \cite{Salio:2002ev}
    \item WT4 represnts ``No-Chaco'' jet events \cite{Vera:2006ib}
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{weather-type-composite-alt.pdf}
  	\caption{
  		Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
      Contours: anomalies of \SI{850}{\hecto\pascal} [contour interval \SI{1e4}{\meter\squared\per\second}]
  	}
    \label{fig:weather-types}
  \end{figure}
  \end{mdframed}
% \item S2S forecasts issued by ECMWF \cite{Vitart2016}
\end{block}

>>>>

\begin{block}{Observations and Weather Types}
  Observations come from:
  \begin{itemize}
    \item Rainfall: CPC Global Unified \cite{xie2010cpc}
    \item Atmosphere: NCAR-NCEP Reanalysis II \cite{Kanamitsu:2002kk}
  \end{itemize}
  We use weather typing \cite{Munoz2015} to represent daily circulation patterns:
  \begin{enumerate}
    \item Calculate streamfunction $\Psi$ from meridional and zonal wind \cite{Dawson:2016ge}
    \item Project \SI{850}{\hecto\pascal} streamfunction onto leading 4 EOFs
    \item $K$-means clustering using classifiability index \cite{Michelangeli1995} to generate single weather type for each day
  \end{enumerate}
  Weather typing simplifies dynamics of daily rainfall but \emph{facilitates analysis of sequences of daily weather patterns}.
  They are associated with patterns that have been well described in the literature; particularly relevant are:
  \begin{itemize}
    \item WT1 represents ``Chaco'' jet event \cite{Salio:2002ev}
    \item WT4 represnts ``No-Chaco'' jet events \cite{Vera:2006ib}
  \end{itemize}
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.925\textwidth]{weather-type-composite-alt.pdf}
  	\caption{
  		Colors: anomalies of rainfall associated with each weather type [\si{\milli\meter\per\day}].
      Contours: anomalies of \SI{850}{\hecto\pascal} streamfunction [contour interval \SI{1e4}{\meter\squared\per\second}]
  	}
    \label{fig:weather-types}
  \end{figure}
  \end{mdframed}
% \item S2S forecasts issued by ECMWF \cite{Vitart2016}
\end{block}

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-model-forecasts.tex
possible move to final poster version

\begin{block}{S2S  Model Forecasts}
  \begin{mdframed}
  \begin{figure}
    \caption{
      Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
      Time series of CPC daily mean precipitation over the same area is plotted with $y$-axis inverted.
  	}
    \noindent\includegraphics[width=0.925\textwidth]{chiclet-s2s-area-averaged.pdf}
  	\label{fig:chiclet}
  \end{figure}
  \end{mdframed}

  \Cref{fig:chiclet} uses a Chiclet diagram to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
  At times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
  At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
  At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.

@@

  In general, \cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
  The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes,.

>>>>

\begin{block}{S2S  Model Forecasts}
  \begin{mdframed}
  \begin{figure}
    \caption{
      Chiclet diagram of ensemble-mean precipitation anomaly forecasts over the Lower Paraguay River Basin from ECMWF S2S forecast data, as a function of the forecast target date (horizontal axis) and lead time (vertical axis).
      Time series of daily precipitation over the same area is plotted with $y$-axis inverted.
  	}
    \noindent\includegraphics[width=0.925\textwidth]{chiclet-s2s-area-averaged.pdf}
  	\label{fig:chiclet}
  \end{figure}
  \end{mdframed}

  \Cref{fig:chiclet} uses a Chiclet diagram \cite{Carbin:2016fx} to visualize, as a function of lead time, the time evolution of the uncorrected, ensemble-mean rainfall anomaly forecast, spatially averaged over the Lower Paraguay River Basin.
  At lead times greater than about two weeks, the ensemble-mean forecast is slightly wetter than climatology.
  At weather timescales (less than one week), the ensemble-mean successfully predicts the timing and amplitude of the area-averaged rainfall.
  At intermediate timescales, the model successfully forecast the strongest breaks and pauses in the rainfall, such as the heavy rainfall during December 2015 and the dry period during mid-January 2016.

  \end{block}
  \begin{block}{Model Output Statistics}

@@

  \Cref{fig:subs-prob-fcst} indicates that better forecasts are obtained when both magnitude and spatial corrections are performed (PCR and CCA).
  The enhanced skill is achieved through the spatial corrections via the EOF-based regressions, which -- in contrast with the extended logistic models -- use information from multiple grid-boxes,.

### 2017-12-02 reports/poster-agu-2017/blocks/s2s-sallj.tex
possible move to final poster version

  During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
  This is consistent with previous studies \cite[i.e.][]{Velasco1987} which find that the intensity and exact location of the anomalous low-level anticylonic anomaly over central Brazil is relevant for the precise impact of ENSO events.

@@

  While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the occurrence of WT4, the ``No-Chaco'' jet event.
  Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region \cite{Vera:2006ib,Barreiro:2017ct,Lima:2017hw}.
  A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.

>>>>

  During El Ni\~{n}o years such as 2015-16, weather type 1 -- related to Chaco jet events -- occurs more frequently for almost all MJO phases, and particularly during phases 1, 2, and 5, consistent with observations of the year under study.
  This agreesß with previous studies \cite[i.e.][]{Velasco1987} which find that the intensity and exact location of the anomalous low-level anticylonic anomaly over central Brazil is relevant for the precise impact of ENSO events.

@@

  While the occurrence of WT1 during NDJF 2015-16 is well explained by ENSO and MJO variability, these features alone do not explain the occurrence of WT4, the ``No-Chaco'' jet event.
  Previous studies emphasize the importance of Pacific-Atlantic interaction for forecasting climate effects in this region \cite{Vera:2006ib,Barreiro:2017ct,Lima:2017hw,Munoz2015}.
  A persistent SST dipole in the central southern Atlantic Ocean favors the occurrence of WT4 by blocking transient extratropical wave activity coming from the Pacific, facilitating transitions from Chaco jet events (WT 1) to No-Chaco jet events (WT 4) via enhanced low-level wind circulation from southern Brazil towards the Atlantic, and back to north-east Brazil and the Amazon (see \cref{fig:chaco-nochaco}) due to land-sea temperature contrasts.
